---
title: "another test"
author: "Ethan Shen"
date: "11/20/2019"
output: html_document
---

```{r}
library(tidyverse) 
```

```{r}
metStandards <- read.csv("met_standards.csv")
metStandards <- metStandards %>%
  filter(Subgroup == "All Students", Grade.Span == "09-12") %>%
  select(School.Name, State.Board.Region, School.Growth.Status, School.Growth.Index.Score)
metStandards[] <- lapply(metStandards, function(x) if(is.factor(x)) as.character(x) else x)

act <- read.csv("act_data.csv")
act <- act %>%
  filter(Subgroup == "All Students", Grade.Span == "09-12", ACT.Subtest.or.Composite == "ACT composite score of 17 or higher") %>%
  select(School.Name, School.Code, Percent.Meeting.Benchmark.or.Standard, Missed.Days.due.to.Hurricane.Florence)
act[] <- lapply(act, function(x) if(is.factor(x)) as.character(x) else x)

SCHOOLINFO <- inner_join(act, metStandards, by = "School.Name")
SCHOOLINFO[] <- lapply(SCHOOLINFO, function(x) if(is.factor(x)) as.character(x) else x)

SCHOOLINFO <- SCHOOLINFO %>%
  rename(school_name = School.Name,
         school_code = School.Code)
```


```{r}
school_data <- read.csv("school_data.csv")

df_1 <- df %>%
  filter(!is.na(grade_span),
         grade_span == "09-12") %>%
  select(school_code, school_name, sbe_region, title_1, missed_days, subgroup, subject, den, total_pct, grade_span)

school_data <- school_data %>%
  filter(!is.na(grade_span),
         grade_span == "09-12") 

df_2 <- reshape(df_1, idvar = "school_name", timevar = "subgroup", direction = "wide")

df_2

df_2$sbe

colnames(df_2)
colnames(df)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

Covariates
- Region (sbe_region)
- title 1 status
- Missed Days (?)
- Class Size
- Percentage Subgroup

- Percent ACT score met standard
ACT <- school_data %>%
  filter(subject == "ACT") 

ACT <- ACT %>%
  filter(school_name != "Dare Learning Academy")

# Title 1 Status
title_1.df <- ACT %>%
  distinct(school_code, school_name, title_1) 

title_1.df[] <- lapply(title_1.df, function(x) if(is.factor(x)) as.character(x) else x)
```

```{r}
#FOR RACE
school.codes <- unique(ACT$school_code)

racedata <- data.frame()
for (code in school.codes){
  
  filteredForCode <- ACT %>%
    filter(school_code == code)
  
  ALL <- filteredForCode %>%
    filter(subgroup == "ALL")
  
  total_number = ALL$den
  sname = ALL$school_name
  
  raceWhite <- filteredForCode %>%
    filter(subgroup == "WHTE")
  raceHispanic <- filteredForCode %>%
    filter(subgroup == "HISP")
  raceBlack <- filteredForCode %>%
    filter(subgroup == "BLCK")
  
  if (nrow(raceWhite) == 1){
    white_number = raceWhite$den
    percent_white = white_number/total_number
  }
  else if (nrow(raceWhite) == 0 && nrow(raceHispanic) == 1 && nrow(raceBlack) == 1){
    hispanic_number = raceHispanic$den
    black_number = raceBlack$den
    minority_number = hispanic_number + black_number
    percent_white = 1 - (minority_number/total_number)
    
  }
  else if (nrow(raceWhite) == 0 && nrow(raceHispanic) == 0 && nrow(raceBlack) == 1){
    black_number = raceBlack$den
    percent_white = 1 - (black_number/total_number)
    
  }
  
  temp.df = data.frame(school_code = code, school_name = sname, 
                       percent_white = percent_white)
  racedata <- rbind(racedata, temp.df)
  racedata[] <- lapply(racedata, function(x) if(is.factor(x)) as.character(x) else x)
}
```

```{r}
#GENDER
names <- unique(ACT$school_code)
genderdata <- data.frame()
for (code in school.codes){
  ALL <- ACT %>%
    filter(school_code == code,
           subgroup == "ALL")
  sname = ALL$school_name
  total_number = ALL$den
  
  #computing stuff for gender
  
  gender <- ACT %>%
    filter(school_code == code,
           subgroup == "MALE" | subgroup == "FEM")
  
  if (nrow(gender) == 0){
    next
  }
  if (nrow(gender) == 2){
    maleInfo <- gender %>%
      filter(subgroup == "MALE")
    
    male_number = maleInfo$den
    percent_male = male_number/total_number
    
  }
  else if (gender$subgroup == "FEM") {
    fem_number = gender$den
    percent_female = fem_number/total_number
    percent_male = 1 - percent_female
    
  }
  
  else if (gender$subgroup == "MALE") {
    male_number = gender$den
    percent_male = male_number/total_number
    
  }
  else{
    next 
  }
  
  temp.df2 = data.frame(school_code = code, school_name = sname, 
                        percent_male = percent_male)
  genderdata <- rbind(genderdata, temp.df2)
  genderdata[] <- lapply(genderdata, function(x) if(is.factor(x)) as.character(x) else x)
}
```

```{r}
# Economically Disadvantaged
ed.data <- data.frame()
for (code in school.codes){
  
  filteredForCode <- ACT %>%
    filter(school_code == code)
  
  ALL <- filteredForCode %>%
    filter(subgroup == "ALL")
  
  total_number = ALL$den
  sname = ALL$school_name
  
  economic.status <- filteredForCode %>%
    filter(subgroup == "EDS" | subgroup == "NEDS")
  
  if (nrow(economic.status) == 0){
    next
  }
  
  if (nrow(economic.status) == 2){
    disadvantaged <- economic.status %>%
      filter(subgroup == "EDS")
    
    eds_number = disadvantaged$den
    percent_eds = eds_number/total_number
    
  }
  
  else if (economic.status$subgroup == "NEDS") {
    fem_number = economic.status$den
    percent_neds = percent_neds/total_number
    percent_male = 1 - percent_neds
    
  }
  
  else if (economic.status$subgroup == "EDS") {
    eds_number = economic.status$den
    percent_eds = eds_number/total_number
    
  }
  else{
    next 
  }
  
  temp.df3 = data.frame(school_code = code, school_name = sname, 
                        percent_eds = percent_eds)
  ed.data <- rbind(ed.data, temp.df3)
  ed.data[] <- lapply(ed.data, function(x) if(is.factor(x)) as.character(x) else x)
}
```

```{r}
# Disabilities 
disability.data <- data.frame()
for (code in school.codes){
  
  filteredForCode <- ACT %>%
    filter(school_code == code)
  
  ALL <- filteredForCode %>%
    filter(subgroup == "ALL")
  
  total_number = ALL$den
  sname = ALL$school_name
  
  disability.status <- filteredForCode %>%
    filter(subgroup == "SWD" | subgroup == "NSWD")
  
  if (nrow(disability.status) == 0){
    next
  }
  
  if (nrow(disability.status) == 2){
    disability <- disability.status %>%
      filter(subgroup == "SWD")
    swd_number = disability$den
    percent_swd = swd_number/total_number
    
  }
  
  else if (disability.status$subgroup == "NSWD") {
    nswd_number = disability.status$den
    percent_nswd = nswd_number/total_number
    percent_swd = 1 - percent_nswd
    
  }
  
  else if (disability.status$subgroup == "SWD") {
    swd_number = disability.status$den
    percent_swd = swd_number/total_number
    
  }
  else{
    next 
  }
  
  temp.df4 = data.frame(school_code = code, 
                        school_name = sname, 
                        percent_swd = percent_swd)
  
  disability.data <- rbind(disability.data, temp.df4)
  disability.data[] <- lapply(disability.data, function(x) if(is.factor(x)) as.character(x) else x)
}
```


```{r, warning = FALSE}
race.gender <- merge(racedata, genderdata, by = c("school_code", "school_name"), all = TRUE) 
race.gender.ed <- merge(race.gender, ed.data, by = c("school_code", "school_name"), all = TRUE) # %>%
#filter(is.na(percent_white) | is.na(percent_male) | is.na(percent_eds))

race.gender.ed.disability <- merge(race.gender.ed, disability.data, by = c("school_code", "school_name"), all = TRUE)

race.gender.ed.disability.title_1 <- merge(race.gender.ed.disability, title_1.df, by = c("school_code", "school_name"), all = TRUE)

final_school_data <- merge(race.gender.ed.disability.title_1, SCHOOLINFO, by = c("school_code", "school_name"), all = TRUE)

final_school_data <- final_school_data %>%
  rename(percent_meeting = Percent.Meeting.Benchmark.or.Standard,
         missed_school_days = Missed.Days.due.to.Hurricane.Florence,
         region = State.Board.Region,
         growth_status = School.Growth.Status,
         growth_index_score = School.Growth.Index.Score) %>%
  mutate(title_1 = case_when(
    title_1 == "" ~ "N",
    TRUE ~ as.character(title_1)),
    percent_meeting = as.numeric(percent_meeting))
```

```{r}
# should we mean center percents? 
model1 = lm(data = final_school_data, growth_index_score ~ .- growth_status - school_code - school_name)

summary(model1)


summary(final_school_data$growth_index_score)
```

```{r}
ggplot(data = final_school_data %>% filter(!is.na(region)), aes(x = percent_white)) + 
  geom_histogram() + 
  facet_wrap(~region)
```

```{r}
avg <- final_school_data %>%
  group_by(region) %>%
  mutate(mean_percent_white = mean(percent_white, na.rm=TRUE),
         mean_percent_male = mean(percent_male, na.rm=TRUE),
         mean_percent_eds = mean(percent_eds, na.rm=TRUE),
         mean_percent_swd = mean(percent_swd, na.rm=TRUE))

avg %>%
  distinct(region, mean_percent_white) %>%
  filter(!is.na(region)) %>%
  spread(region, mean_percent_white)
```

