---
title: "Untitled"
author: "Ethan Shen"
date: "11/6/2019"
output: html_document
---

```{r}
met_standards
library(tidyverse)
```
df <- read.csv("asdf.csv")



```{r}
df <- read.csv("asdf.csv")

df <- df %>%
   filter(!is.na(grade_span),
          grade_span == "09-12") 
reshape(df, idvar = "school_name", timevar = "subgroup", direction = "wide")


a <- df %>%
   filter(subject == "ACT" |  
             subject == "ACTEN" |  subject == "ACTMA" | subject == "ACTRD" | subject == "ACTSC" |  subject == "ACTWR") %>%
   na.omit()

unique(a$sbe_region)

a[, ] <- lapply(a[, ], as.character)
a %>%
   select(-reporting_year, -lea_code, -lea_name, -school_code, -grade_span, -notprof_pct, -lev1_pct, -lev2_pct, -lev3_pct, -lev4_pct, -lev5_pct, -glp_pct, -ccr_pct) %>%
   spread(key = "subgroup", value = "total_pct") 
b <- a  %>%
   spread(key = "subgroup", value = "total_pct") 
```


```{r}
b %>%
   filter(school_name == "Eastern Alamance High",  subject == "ACT")

lm(lea_code ~ as.numeric(AIG), data = b)

test.df <- data.frame()
try=function(oneSchoolInfo) {
   
   total_number <- oneSchoolInfo %>%
      filter(
         subgroup == "ALL") %>%
      select(den)
   percent_female <- oneSchoolInfo %>%
      filter(
         subgroup == "FEM") %>%
      select(den)
   percent_female = percent_female/total_number
   dat = data.frame(total = total_number, percent_fem = percent_female)
   return(rbind(test.df, dat))
}

try(test)
```


```{r}
for (one in names) {
   total_number <- test %>%
      filter(school_name == one,
             subgroup == "ALL") %>%
      select(den)
   percent_female <- test %>%
      filter(school_name == one,
             subgroup == "FEM") %>%
      select(den)
   percent_female = percent_female/total_number
   l = data.frame(schooltotal = total_number, percent_fem = percent_female)
   return(l)
}

try(test)

df_just_act <- df %>%
   filter(subject == "ACT")
df_just_act

test <- df_just_act %>%
   filter(school_name == 'Eastern Alamance High' | school_name == "Graham High")

test[, ] <- lapply(test[, ], as.character)

names <- unique(test$school_name)  
unique(test$subgroup)  
o <- data.frame()
for (one in names) {
   ALL <- test %>%
      filter(school_name == one,
             subgroup == "ALL") 
   total_number = ALL$den
   
   FEM <- test %>%
      filter(school_name == one,
             subgroup == "FEM")
   fem_number = FEM$den
   
   MALE <- test %>%
      filter(school_name == one,
             subgroup == "MALE")
   male_number = MALE$den
   
   BLCK <- test %>%
      filter(school_name == one,
             subgroup == "BLCK")
   blck_number = BLCK$den
   
   HISP <- test %>%
      filter(school_name == one,
             subgroup == "HISP")
   hisp_number = HISP$den
   
   MULT <- test %>%
      filter(school_name == one,
             subgroup == "MULT")
   mult_number = MULT$den
   
   WHTE <- test %>%
      filter(school_name == one,
             subgroup == "WHTE")
   whte_number = WHTE$den
   
   EDS <- test %>%
      filter(school_name == one,
             subgroup == "EDS")
   eds_number = EDS$den
   
   SWD <- test %>%
      filter(school_name == one,
             subgroup == "SWD")
   swd_number = SWD$den
   
   AIG <- test %>%
      filter(school_name == one,
             subgroup == "AIG")
   aig_number = AIG$den
   
   NEDS <- test %>%
      filter(school_name == one,
             subgroup == "NEDS")
   neds_number = NEDS$den
   
   NSWD <- test %>%
      filter(school_name == one,
             subgroup == "NSWD")
   nswd_number = NSWD$den
   
   ELS <- test %>%
      filter(school_name == one,
             subgroup == "ELS")
   els_number = ELS$den
   
   percent_female = fem_number/total_number
   percent_male = male_number/total_number
   percent_black = blck_number/total_number
   percent_hispanic = hisp_number/total_number
   percent_mult = mult_number/total_number
   percent_white = whte_number/total_number
   percent_eds = eds_number/total_number
   percent_swd = swd_number/total_number
   percent_aig = aig_number/total_number
   percent_neds = neds_number/total_number
   percent_nswd = nswd_number/total_number
   percent_els = els_number/total_number
   
   df = data.frame(school_name = one, 
                   percent_female = percent_female,
                   percent_male = percent_male,
                   # percent_black = percent_black,
                   # percent_hispanic = percent_hispanic,
                   # percent_mult = percent_mult,
                   # percent_white = percent_white,
                   # percent_eds = percent_eds,
                   # percent_swd = percent_swd,
                   # percent_aig = percent_aig,
                   # percent_neds = percent_neds,
                   percent_nswd = percent_nswd,
                   percent_els = percent_els)
                 
   o <- rbind(o, df)
                   
}
```


```{r}
d <- tibble(hits = 1, runs = 2, o_hits = 5, o_runs = 9)
d2 <- d %>% gather(key = "col_name", value = "value", everything())
d3 <- d2 %>% separate(col_name, into = c("team", "metric"), sep = "_", fill = "left")
d3$team <- ifelse(is.na(d3$team), "a", "b")
d4 <- d3 %>% group_by(metric) %>% spread(key = "team", value = "value")
d5 <- d4 %>% mutate(diff = a - b)
```


set hierarchy on school 

### Each row is a school here  
```{r}
df <- df %>%
filter(!is.na(grade_span),
grade_span == "09-12", subject == "ACT" |
subject == "ACTEN" |subject == "ACTMA" | subject == "ACTRD" | subject == "ACTSC" |subject == "ACTWR")

df

reshape(df, idvar = "school_name", timevar = "subgroup", direction = "wide")
```


Response: Growth score

Covariates
- Region (sbe_region)
- title 1 status
- Missed Days (?)
- School Size
- Percentage Subgroup
- Percent ACT score met standard
