---
title: "Project Modeling"
author: "Ethan Shen, Steven Herrera, Malavi Ravindran, Austin Jia"
date: "11/23/2019"
output: html_document
---

```{r, echo = FALSE, include = FALSE} 
# Installing packages
pkgTest <- function(x) {
  if (!require(x,character.only = TRUE)) {
    install.packages(x,dep=TRUE) 
  }
}
pkgs <- c("tidyverse", "ggplot2", "ggpubr", "cowplot", "reshape2", "caTools", "MASS", "psych")

for (pkg in pkgs) {
  pkgTest(pkg)
}
```

```{r}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(reshape2)
library(caTools) 
library(MASS)
require(foreign)
require(Hmisc)
require(reshape2)
library(e1071)
library(caret)
library(psych)
library(randomForest)
```

```{r message = FALSE}
school_df_final <- read_csv("school_df_final.csv")
school_df_final <- school_df_final %>%
  na.omit()
```

### Visualizations

```{r}
yes = c("#66c2a5",
        "#fc8d62",
        "#8da0cb",
        "#e78ac3")


ya = c(
  "#8dd3c7",
  "#ffffb3",
  "#bebada",
  "#fb8072"
)

heh = c(
  "#1f78b4",
  "#fdbf6f",
  "#33a02c",
  "#fb9a99"
)
avg.dem.dist <- school_df_final %>%
  group_by(region) %>%
  mutate(mean_percent_white = mean(percent_white, na.rm = TRUE),
         mean_percent_male = mean(percent_male, na.rm = TRUE),
         mean_percent_eds = mean(percent_eds, na.rm = TRUE),
         mean_percent_swd = mean(percent_swd, na.rm = TRUE)) %>%
  distinct(region, 
           mean_percent_white,
           mean_percent_male,
           mean_percent_eds,
           mean_percent_swd) %>%
  gather(variable, value, -region) 

ggplot(avg.dem.dist, aes(x = region, y = value * 100, fill = reorder(variable, -value))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(name = "Demographic Label", 
                    labels = c("White", "Male", "Economically Disadvantaged", "Students with Disabilities"),
                    values = heh#c("#a1dab4", "#41b6c4", "#225ea8", "#253494")
  ) + 
  #  scale_fill_discrete(name = "Demographic Label", labels = c("White", "Male", "EDS", "SWD")) + 
  
  labs(title = "Average Demographic Distribution by Region", 
       x = "Region", 
       y = "Percent") + 
  theme_minimal() + 
  rotate_x_text(angle = 50)
```

```{r}
#TRY BOX COX. SCREEn. 

col.pal = c("#8dd3c7",
            "#ffffb3",
            "#bebada",
            "#fb8072",
            "#80b1d3",
            "#fdb462",
            "#b3de69",
            "#fccde5")


col.pal.2 = c('#66c2a5',
              '#fc8d62',
              '#8da0cb',
              '#e78ac3',
              '#a6d854',
              '#ffd92f',
              '#e5c494',
              '#b3b3b3')

ggplot(school_df_final, aes(x = growth_index_score, fill = region)) + 
  geom_histogram(binwidth = 0.9, show.legend = FALSE) + 
  facet_wrap(~region, nrow = 4) + 
  labs(title = "Histogram of Growth Index Score by Region", 
       x = "Growth Index Score", 
       y = "Count") +
  scale_fill_manual(values = col.pal.2) +  
  theme_bw() 
```


```{r}

ggplot(data = school_df_final %>% filter(!is.na(region)), aes(x = percent_white)) + 
  geom_histogram() + 
  facet_wrap(~region)
```


```{r}
avg <- school_df_final %>%
  group_by(region) %>%
  mutate(mean_percent_white = mean(percent_white, na.rm=TRUE),
         mean_percent_male = mean(percent_male, na.rm=TRUE),
         mean_percent_eds = mean(percent_eds, na.rm=TRUE),
         mean_percent_swd = mean(percent_swd, na.rm=TRUE))
avg %>%
  distinct(region, mean_percent_white) %>%
  filter(!is.na(region)) %>%
  spread(region, mean_percent_white)

ACT.info <- school_df_final %>%
  dplyr::select(percent_ACTEN_meeting, 
                percent_ACTMA_meeting,
                percent_ACTRD_meeting,
                percent_ACTSC_meeting,
                percent_ACTWR_meeting,
                percent_ACT_meeting)
cor.ACT <- cor(ACT.info, use = "complete.obs")

get_upper_tri <- function(cormat){
  cormat[upper.tri(cormat)]<- NA
  return(cormat)
}
upper_tri <- get_upper_tri(cor.ACT)

melted_cormat <- melt(upper_tri, na.rm = TRUE)


ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(color = "white") + 
  labs(title = "Correlation Matrix of ACT Subcategories", 
       x = "",
       y = "") + 
  scale_fill_gradient2(low = "#edf8b1", mid = "#7fcdbb", high = "#2c7fb8", 
                       midpoint = (min(cor.ACT) + 1)/2, limit = c(min(cor.ACT),1), space = "Lab", 
                       name="Correlation") + 
  theme_minimal() +
  rotate_x_text(angle = 50) 
```

```{r}
numeric.cols <- school_df_final[ , purrr::map_lgl(school_df_final, is.numeric)]

pairs.panels(numeric.cols[c(ncol(numeric.cols), 1:4)],
                    smooth = FALSE,
                    pch = 1,
                    ellipses = FALSE,
                    hist.col = "#80b1d3")
pairs.panels(numeric.cols[c(ncol(numeric.cols), 5:8)],
                    smooth = FALSE,
                    pch = 1,
                    ellipses = FALSE,
                    hist.col = "#80b1d3")

pairs.panels(numeric.cols[c(ncol(numeric.cols), 9:11)],
                    smooth = FALSE,
                    pch = 1,
                    ellipses = FALSE,
                    hist.col = "#80b1d3")

pairs(numeric.cols[c(ncol(numeric.cols), 5:8)],
      lower.panel = NULL)

pairs(numeric.cols[c(ncol(numeric.cols), 9:11)],
      lower.panel = NULL)

```


### Modeling 

```{r}
#first we omit any remaining NA values from our data
school_df_final<-na.omit(school_df_final) 
school_df_final
```

#### Regression



##### Linear
```{r}
#linear model with selection- first, we are fitting a full model (except for those specific ACT variables we are cutting)
linear.fit1 <- lm(growth_index_score ~ percent_white + percent_male + percent_eds + percent_swd + title_1 + percent_ACT_meeting + missed_school_days + region, data = school_df_final)
summary(linear.fit1)

```

Our full model does not perform well (very low adjusted R squared). We will try performing stepwise selection on the model to see if we can improve performance.


```{r}
#linear model with stepwise selection
linear.fit2 <- stepAIC(linear.fit1, direction = "both", trace = FALSE)
summary(linear.fit2)
```

It is clear that race, region, and ACT performance are strong indicators of performance. Are any interactions significant?

```{r}
#different linear model with interactions
linear.fit3 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*percent_ACT_meeting +  percent_white*region
                  + region*percent_ACT_meeting, data = school_df_final)

linear.fit4 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*percent_ACT_meeting +  percent_white*region, data = school_df_final)

linear.fit5 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*region
                  + region*percent_ACT_meeting, data = school_df_final)

linear.fit6 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*percent_ACT_meeting + region*percent_ACT_meeting, data = school_df_final)

linear.fit7 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region +  percent_white*percent_ACT_meeting, data = school_df_final)

linear.fit8 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region +  percent_white*region, data = school_df_final)

linear.fit9 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region +  region*percent_ACT_meeting, data = school_df_final)

AIC(linear.fit2)
AIC(linear.fit3)
AIC(linear.fit4)
AIC(linear.fit5)
AIC(linear.fit6)
AIC(linear.fit7)
AIC(linear.fit8)
AIC(linear.fit9)

```

So, the model with the lowest AIC is the following:

growth_index_score ~ percent_white + percent_ACT_meeting  + region +  region*percent_ACT_meeting

```{r}
summary(linear.fit9)
```

Visualizing interactions

```{r}
ggplot(data=school_df_final, aes(x=percent_ACT_meeting ,y=growth_index_score, color= region)) + 
  geom_smooth(se = F, method = "lm")

```

```{r}
plot(linear.fit9)
```




```{r}
school_df_final$region <- as.factor(school_df_final$region)
```

```{r}
gam1 <- mgcv::gam(growth_index_score ~ s(percent_ACT_meeting) + s(percent_white) + region, data = school_df_final)
gam2 <- mgcv::gam(growth_index_score ~ percent_ACT_meeting + s(percent_white) + region, data = school_df_final)
gam3 <-  mgcv::gam(growth_index_score ~ s(percent_ACT_meeting) + percent_white + region, data = school_df_final)
gam4 <-  mgcv::gam(growth_index_score ~ s(percent_ACT_meeting) + percent_white + region + region*percent_ACT_meeting, data = school_df_final)
gam5 <-  mgcv::gam(growth_index_score ~ s(percent_ACT_meeting) + s(percent_white) + region + region*percent_ACT_meeting, data = school_df_final)
AIC(gam1)
AIC(gam2)
AIC(gam3)
AIC(gam4)
AIC(gam5)

```

```{r}
plot(gam3, all.terms = TRUE, seWithMean = TRUE)

#EDA Boxplot
ggplot(school_df_final,aes(x=region,y=growth_index_score,fill=region))+
  geom_boxplot(position=position_dodge(1))+
  ggtitle("Growth Index Score Distribution by Region") +
  ylab("Growth Index Score") +
  scale_x_discrete("Region", drop=FALSE) +
  theme(axis.text.x=element_text(angle=45,hjust=1)) 
```


```{r}
school_df_final$growth_status <- factor(school_df_final$growth_status, levels =c("Not Met","Met", "Exceeded"), ordered=TRUE)
school_df_final
```


```{r}
ord.logistic <- polr(growth_status ~ percent_white + percent_ACT_meeting +  region + region*percent_ACT_meeting, data = school_df_final, Hess=TRUE)
summary(ord.logistic)
```



```{r}
testing_dropped_columns <- school_df_final %>%
  dplyr::select(school_code, school_name, region, percent_white, percent_ACT_meeting, growth_status)
testing.with.predicted.probs <- cbind(testing_dropped_columns, predict(ord.logistic, school_df_final, type = "probs"))
reshaped.predictions <- melt(testing.with.predicted.probs, id.vars = c("school_code", "school_name", "region", "percent_white", "percent_ACT_meeting", "growth_status" ),
  variable.name = "Level", value.name="Probability")
## view first few rows
head(reshaped.predictions)
reshaped.predictions %>%
  dplyr::select(region, Level, Probability) %>%
  group_by(region, Level) %>%
  mutate(n = n()) %>%
  mutate(s = sum(Probability))
```



```{r}
ggplot(reshaped.predictions, aes(x = percent_white, y = Probability, colour = Level)) +
  geom_smooth()


ggplot(reshaped.predictions, aes(x = percent_ACT_meeting, y = Probability, colour = Level)) +
  geom_smooth()

ggplot(reshaped.predictions, aes(x = region, y = Probability, fill = Level)) +
  geom_bar(stat = "identity", position = "fill") + theme_bw() + 
  rotate_x_text(angle = 50)


```






