---
title: "Project Modeling"
author: "Ethan Shen, Steven Herrera, Malavi Ravindran, Austin Jia"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmdformats::readthedown
---

# Loading Packages 

```{r, echo = FALSE, include = FALSE}
# Installing packages
pkgTest <- function(x) {
  if (!require(x,character.only = TRUE)) {
    install.packages(x,dep=TRUE) 
  }
}
pkgs <- c("tidyverse", "ggplot2", "ggpubr", "cowplot", "reshape2", "caTools", "MASS", "psych")

for (pkg in pkgs) {
  pkgTest(pkg)
}
```

```{r message = FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(reshape2)
library(caTools) 
library(MASS)
require(foreign)
require(Hmisc)
require(reshape2)
library(e1071)
library(caret)
library(psych)
library(randomForest)
library(gridExtra)
library(grid)
library(jtools)


ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = 'center')
```

# Reading Data 

```{r message = FALSE}
school_df_final <- read_csv("school_df_final.csv")
school_df_final <- school_df_final %>%
  na.omit()
```

# EDA

## Response Variable 

```{r}
col.pal.2 = c('#66c2a5',
            '#fc8d62',
            '#8da0cb',
            '#e78ac3',
            '#a6d854',
            '#ffd92f',
            '#e5c494',
            '#b3b3b3')

p1 <- ggplot(school_df_final, aes(x = growth_index_score)) + 
   geom_histogram(binwidth = 0.9, show.legend = FALSE, col = "black", fill = "#74a9cf") +
   labs(y = "Count")+ 
   theme_bw() + 
   theme(axis.title.x=element_blank(),
         axis.title.y=element_blank())

p2 <- ggplot(school_df_final, aes(x = growth_index_score, fill = region)) + 
   geom_histogram(binwidth = 0.9, show.legend = FALSE) + 
   facet_wrap(~region, nrow = 4) + 
   labs(subtitle = "by Region",
        x = "Growth Index Score", 
        y = "Count") +
   scale_fill_manual(values = col.pal.2) + 
   theme_bw() +
   theme(axis.title.x=element_blank(),
         axis.title.y=element_blank())


 grid.arrange(arrangeGrob(plot_grid(p1,p2,ncol=2), 
                          top = textGrob("Histograms of Growth Index Score",gp=gpar(fontsize=15)), 
                          left = textGrob("Count",gp=gpar(fontsize=11), rot = 90),
                          bottom = textGrob("Growth Index Score",gp=gpar(fontsize=11))))
```

We first examine one of our response variable, and see that the growth score index as a whole and across the regions is approximately normally distributed, which means we do not have to perform transformations on our response variable.

## Demographics Across Regions

```{r}
col.pal = c(
  "#1f78b4",
  "#fdbf6f",
  "#33a02c",
  "#fb9a99"
)
avg.dem.dist <- school_df_final %>%
  group_by(region) %>%
  mutate(mean_percent_white = mean(percent_white, na.rm = TRUE),
         mean_percent_male = mean(percent_male, na.rm = TRUE),
         mean_percent_eds = mean(percent_eds, na.rm = TRUE),
         mean_percent_swd = mean(percent_swd, na.rm = TRUE)) %>%
  distinct(region, 
           mean_percent_white,
           mean_percent_male,
           mean_percent_eds,
           mean_percent_swd) %>%
  gather(variable, value, -region) 

ggplot(avg.dem.dist, aes(x = region, y = value, fill = reorder(variable, -value))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  scale_fill_manual(name = "Demographic Label", 
                    labels = c("White", "Male", "Economically Disadvantaged", "Students with Disabilities"),
                    values = col.pal) + 
  labs(title = "Average Demographic Distribution by Region", 
       x = "Region", 
       y = "Percent") + 
  theme_minimal() + 
  rotate_x_text(angle = 50)
```

This plot shows us some of the different average demographic percentages across the regions. We see that the percent of students with disabilities and percent of male students is about the same across all regoins. However, the other two percentages differ quite a bit across regions, with the Northwest and Western regions having over an average of 70% white students and the Sandhill region having almost 60% of economically disadvantaged students.

## Pairs Plots 

```{r}
numeric.cols <- school_df_final[ , purrr::map_lgl(school_df_final, is.numeric)]

pairs.panels(numeric.cols[c(ncol(numeric.cols), 1:4)],
             smooth = FALSE,
             pch = 1,
             ellipses = FALSE,
             hist.col = "#80b1d3")
pairs.panels(numeric.cols[c(ncol(numeric.cols), 5:8)],
             smooth = FALSE,
             pch = 1,
             ellipses = FALSE,
             hist.col = "#80b1d3")

pairs.panels(numeric.cols[c(ncol(numeric.cols), 9:11)],
             smooth = FALSE,
             pch = 1,
             ellipses = FALSE,
             hist.col = "#80b1d3")
```

We then plotted pairs plots, which also show the univariate distribution of the variable, and provide a scatterplot and correlation of the bivariate relationships. For the most part, multicollinearity is not a big issue; however, some predictors (namely ones associated with the ACT subsections) are highly correlated. We dive deeper into those predictors in the following correlation matrix. 

## Correlation Matrix 

```{r}
avg <- school_df_final %>%
  group_by(region) %>%
  mutate(mean_percent_white = mean(percent_white, na.rm=TRUE),
         mean_percent_male = mean(percent_male, na.rm=TRUE),
         mean_percent_eds = mean(percent_eds, na.rm=TRUE),
         mean_percent_swd = mean(percent_swd, na.rm=TRUE))
avg %>%
  distinct(region, mean_percent_white) %>%
  filter(!is.na(region)) %>%
  spread(region, mean_percent_white)

ACT.info <- school_df_final %>%
  dplyr::select(percent_ACTEN_meeting, 
                percent_ACTMA_meeting,
                percent_ACTRD_meeting,
                percent_ACTSC_meeting,
                percent_ACTWR_meeting,
                percent_ACT_meeting)
cor.ACT <- cor(ACT.info, use = "complete.obs")

get_upper_tri <- function(cormat){
  cormat[upper.tri(cormat)]<- NA
  return(cormat)
}
upper_tri <- get_upper_tri(cor.ACT)

melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile(color = "white") + 
  labs(title = "Correlation Matrix of ACT Subcategories", 
       x = "",
       y = "") + 
  scale_fill_gradient2(low = "#edf8b1", mid = "#7fcdbb", high = "#2c7fb8", 
                       midpoint = (min(cor.ACT) + 1)/2, limit = c(min(cor.ACT),1), space = "Lab", 
                       name="Correlation") + 
  theme_minimal() +
  rotate_x_text(angle = 50) 
```

We see here that the ACT subcategory predictors are all very highly correlated. Thus, when moving forward with the modeling, we will only keep the overall ACT benchmark percentage (`percent_ACT_meeting`).

# Regression Modeling 

```{r}
#splitting data into training and testing

school_df_final$title_1 <- as.factor(school_df_final$title_1)
school_df_final$region <- as.factor(school_df_final$region)
school_df_final$growth_status <- as.factor(school_df_final$growth_status)

# set.seed(123)   #  set seed to ensure you always have same random numbers generated
# rows <- sample(nrow(school_df_final))
# school_df_final <- school_df_final[rows, ]
# sample = sample.split(school_df_final,SplitRatio = 0.75) # splits the data in the ratio mentioned in SplitRatio. After splitting marks these rows as logical TRUE and the the remaining are marked as logical FALSE
# training =subset(school_df_final,sample ==TRUE) # creates a training dataset named train1 with rows which are marked as TRUE
# testing=subset(school_df_final, sample==FALSE)
# #first we omit any remaining NA values from our data
# school_df_final<-na.omit(school_df_final) 
# school_df_final
#first we omit any remaining NA values from our data
# school_df_final<-na.omit(school_df_final) 
# school_df_final
```

## Linear
```{r}
#linear model with selection- first, we are fitting a full model (except for those specific ACT variables we are cutting)
linear.fit1 <- lm(growth_index_score ~ percent_white + percent_male + percent_eds + percent_swd + title_1 + percent_ACT_meeting + missed_school_days + region, data = school_df_final)
summary(linear.fit1)

```

Our full model does not perform well (very low adjusted R squared). We will try performing stepwise selection on the model to see if we can improve performance.


```{r}
#linear model with stepwise selection
linear.fit2 <- stepAIC(linear.fit1, direction = "both", trace = FALSE)
summary(linear.fit2)
```

It is clear that race, region, and ACT performance are strong indicators of performance. Are any interactions significant?

```{r}
#different linear model with interactions
linear.fit3 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*percent_ACT_meeting +  percent_white*region
                  + region*percent_ACT_meeting, data = school_df_final)

linear.fit4 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*percent_ACT_meeting +  percent_white*region, data = school_df_final)

linear.fit5 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*region
                  + region*percent_ACT_meeting, data = school_df_final)

linear.fit6 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region + percent_white*percent_ACT_meeting + region*percent_ACT_meeting, data = school_df_final)

linear.fit7 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region +  percent_white*percent_ACT_meeting, data = school_df_final)

linear.fit8 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region +  percent_white*region, data = school_df_final)

linear.fit9 <- lm(growth_index_score ~ percent_white + percent_ACT_meeting  + region +  region*percent_ACT_meeting, data = school_df_final)

AIC(linear.fit2)
AIC(linear.fit3)
AIC(linear.fit4)
AIC(linear.fit5)
AIC(linear.fit6)
AIC(linear.fit7)
AIC(linear.fit8)
AIC(linear.fit9)

```

So, the model with the lowest AIC is the following:

growth_index_score ~ percent_white + percent_ACT_meeting  + region +  region*percent_ACT_meeting

```{r}
summary(linear.fit9)
plot_summs(linear.fit9, scale = TRUE, plot.distributions = TRUE)
```

Visualizing interactions

```{r}
ggplot(data=school_df_final, aes(x=percent_ACT_meeting ,y=growth_index_score, color = region)) + 
  geom_smooth(se = F, method = "lm") + 
  labs(title = "Growth Index Score vs. Percent Meeting ACT Benchmark",
       x = "% Meeting ACT Benchmark",
       y = "Growth Index Score") + 
  scale_color_manual(values = col.pal.2, name = "Region")
ggplot(data=school_df_final, aes(x=percent_ACT_meeting ,y=growth_index_score, color= region)) + 
  geom_smooth(se = F, method = "lm") + labs(title = "Growth Index Score vs. Percent Meeting ACT Standards by Region") 

```

```{r}
plot(linear.fit9)
BIC(linear.fit9)
```




```{r}
school_df_final$region <- as.factor(school_df_final$region)
```

```{r}
gam1 <- mgcv::gam(growth_index_score ~ s(percent_ACT_meeting) + s(percent_white) + region, data = school_df_final)
gam2 <- mgcv::gam(growth_index_score ~ percent_ACT_meeting + s(percent_white) + region, data = school_df_final)
gam3 <-  mgcv::gam(growth_index_score ~ s(percent_ACT_meeting) + percent_white + region, data = school_df_final)
gam4 <- mgcv::gam(growth_index_score ~ s(percent_ACT_meeting, by = region) + s(percent_white, by = region) + region, data = school_df_final)
gam5 <- mgcv::gam(growth_index_score ~ s(percent_ACT_meeting, by = region) + percent_white + region, data = school_df_final)
gam6 <- mgcv::gam(growth_index_score ~ s(percent_ACT_meeting, by = region) + s(percent_white) + region, data = school_df_final)


AIC(gam1)
AIC(gam2)
AIC(gam3)
AIC(gam4)
AIC(gam5)
AIC(gam6)

BIC(gam6)

plot(gam6, all.terms = TRUE, seWithMean = TRUE, pages = 1)

```

```{r}
plot(gam3, all.terms = TRUE, seWithMean = TRUE)

#EDA Boxplot
ggplot(school_df_final,aes(x=region,y=growth_index_score,fill=region))+
  geom_boxplot(position=position_dodge(1))+
  ggtitle("Growth Index Score Distribution by Region") +
  ylab("Growth Index Score") +
  scale_x_discrete("Region", drop=FALSE) + 
  scale_fill_manual(values = col.pal.2, name = "Region") + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) 

plot(gam6, all.terms = TRUE, seWithMean = TRUE)
```


```{r}
school_df_final$growth_status <- factor(school_df_final$growth_status, levels =c("Not Met","Met", "Exceeded"), ordered=TRUE)
school_df_final
```


```{r}
ord.logistic <- polr(growth_status ~ percent_white + percent_ACT_meeting +  region + region*percent_ACT_meeting, data = school_df_final, Hess=TRUE)
summary(ord.logistic)
```



```{r}
testing_dropped_columns <- school_df_final %>%
  dplyr::select(school_code, school_name, region, percent_white, percent_ACT_meeting, growth_status)
testing.with.predicted.probs <- cbind(testing_dropped_columns, predict(ord.logistic, school_df_final, type = "probs"))
reshaped.predictions <- melt(testing.with.predicted.probs, id.vars = c("school_code", "school_name", "region", "percent_white", "percent_ACT_meeting", "growth_status" ),
                             variable.name = "Level", value.name="Probability")
## view first few rows
head(reshaped.predictions)
reshaped.predictions %>%
  dplyr::select(region, Level, Probability) %>%
  group_by(region, Level) %>%
  mutate(n = n()) %>%
  mutate(s = sum(Probability))
```



```{r}

ggplot(reshaped.predictions, aes(x = percent_white, y = Probability, colour = Level)) +
  geom_smooth(se = F) + 
  labs(title = "Probability of Meeting Standards by Percent White",
       x = "% White") + 
  theme_bw()

ggplot(reshaped.predictions, aes(x = percent_ACT_meeting, y = Probability, colour = Level)) +
  geom_smooth(se = F) + 
  labs(title = "Probability of Meeting Standards by Percent Meeting ACT Standards",
       x = "% Meeting ACT Benchmark") + theme_bw()

ggplot(reshaped.predictions, aes(x = region, y = Probability, fill = Level)) +
  geom_bar(stat = "identity", position = "fill") + theme_bw() + 
  rotate_x_text(angle = 50) + 
  labs(title = "Probability of Meeting Standards by Region",
       x = "Region")


```






