---
title: "STA 325 Project Proposal"
author: "Ethan Shen, Steven Herrera, Malavi Ravindran, Austin Jia"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: rmdformats::readthedown
---

```{r}
library(tidyverse)
```

# Project Summary 

# Data Description 

```{r}
tweets <- read_csv("tweet_global_warming.csv") %>%
  select(tweet, existence)
head(tweets)
```

## Data Source 

We got the data from https://www.figure-eight.com/data-for-everyone/. 

## Predictors and Response

We will be using the `tweet` column as our predictor. Our response variable will be `existence`. 

## Data Type 

`tweet` is a column that contains various, and is unstructured in the sense that it does not fall into a specific variable category (it's not continuous or discrete). `existence` is a nominal categorical variable. 

## Data Scraping / Wrangling 

We will be using various NLP techniques to transform the `tweet` column, such as determining word frequency (how often a word appears in these tweets), word length, lexical diversity (the number of unique words used in a tweet), and lexical density (the number of unique words divided by the total number of words for each tweet).

# Project Roadmap 
